Differential invariant signatures for planar Lie group transformations with application to images

### 2021-09-14 paper/newerpaper.tex
More minor increments

In order to define differential invariants, it is necessary to identify how the group transformation $\varphi$ acts on derivatives of the function $f$ via the jet space~\citep{OlverEIS}. The $d$-th order jet space $J^d$ consists of the function itself and the set of all partial derivatives of the function up to order $d$. (Note that there are $\left( \begin{array}{c} d+k-1 \\ k \end{array} \right)$ partial derivatives of order $d$.) As the group $\mathcal{G}$ is local, it acts naturally on derivatives of the transformation, and so each individual transformation $\varphi \in \mathcal{G}$ can be prolonged to act on the jet space $J^d$. We can thus prolong the action of the group and consider the action of $\mathcal{G}$ on not just $f$, but all derivatives of $f$ up to order $d$. A $d$-th order differential invariant is then a local scalar function $I: J^{d} (f,\Omega, k) \to \mathbb{R}$ that is invariant under the action of $\mathcal{G}$, i.e., $I(\varphi(f,\Omega, k)) = I(f, \Omega, k) \forall \varphi \in \mathcal{G}$. It can also be useful to consider the maximum order of derivative in the invariant, which we term degree, e.g., $f_{xy}^2 f_x f_y$ has degree 2.

@@

Although there are computational challenges regarding the numerical approximation of derivatives on images, the fact that differential invariants are robust to occlusion, cheap to compute, and well-behaved with respect to small changes in the image, has made them popular for researchers in image analysis~\citep{Florack1993}. Two particular tools have been key to this: Cartan's moving frame~\citep{Cartan35} (see Section~\ref{sec:movingframes}), and which was used to construct image invariants as early as 1994 by \citet{Faugeras1994}, and the concept of an image signature built from a set of invariants~\citep{Calabi1998}. An algorithm to compute differential and joint invariants and determine complete signature sets is described by \citet{Olver2005,Olver2007}. The entire set of invariants is computed using the method of moving frames and then the signature set of a maximal set of functionally independent invariants is identified. A third step is possible, in which the dimension of the signature set is reduced by eliminating those invariants whose values can be computed from the values of derivatives of some subset of the invariants.

@@

So far, the methods that we have seen to construct differential invariants for particular groups are not directly generalisable. It is natural to ask if there is an algorithmic approach that will yield invariants more directly. The answer is a qualified yes. There is a general method, based on Cartan's method of moving frames, that is generally applicable, but finding good solutions with it still requires some ingenuity in general. Olver and co-authors have popularised the approach and made it applicable to object recognition based on outline curves, see~\citet{Olver2005} for an overview, and note particularly the original application paper~\citep{Calabi1998}.  We introduce the method here and then demonstrate its use to find invariants to further groups, but for further technical details see~\citet{OlverCIT}.

>>>>

In order to define differential invariants, it is necessary to identify how the group transformation $\varphi$ acts on derivatives of the function $f$ via the jet space~\citep{Kruglikov2011,OlverEIS}. The $d$-th order jet space $J^d$ consists of the function itself and the set of all partial derivatives of the function up to order $d$. (Note that there are $\left( \begin{array}{c} d+k-1 \\ k \end{array} \right)$ partial derivatives of order $d$.) As the group $\mathcal{G}$ is local, it acts naturally on derivatives of the transformation, and so each individual transformation $\varphi \in \mathcal{G}$ can be prolonged to act on the jet space $J^d$. We can thus prolong the action of the group and consider the action of $\mathcal{G}$ on not just $f$, but all derivatives of $f$ up to order $d$. A $d$-th order differential invariant is then a local scalar function $I: J^{d} (f,\Omega, k) \to \mathbb{R}$ that is invariant under the action of $\mathcal{G}$, i.e., $I(\varphi(f,\Omega, k)) = I(f, \Omega, k) \forall \varphi \in \mathcal{G}$. It can also be useful to consider the maximum order of derivative in the invariant, which we term degree, e.g., $f_{xy}^2 f_x f_y$ has degree 2.

@@

Although there are computational challenges regarding the numerical approximation of derivatives on images, the fact that differential invariants are robust to occlusion, cheap to compute, and well-behaved with respect to small changes in the image, has made them popular for researchers in image analysis~\citep{Florack1993}. Two particular tools have been key to this: Cartan's moving frame~\citep{Cartan35} (see Section~\ref{sec:movingframes}), and which was used to construct image invariants as early as 1994 by \citet{Faugeras1994}, and the concept of an image signature built from a set of invariants~\citep{Calabi1998}. An algorithm to compute differential and joint invariants and determine complete signature sets is described by \citet{Olver2005,Olver2007}
The entire set of invariants is computed using the method of moving frames and then the signature set of a maximal set of functionally independent invariants is identified. A third step is possible, in which the dimension of the signature set is reduced by eliminating those invariants whose values can be computed from the values of derivatives of some subset of the invariants. Further work, including applications to ODEs and the Calculus of Variations can be found in~\cite{Mansfield}.

@@

So far, the methods that we have seen to construct differential invariants for particular groups are not directly generalisable. It is natural to ask if there is an algorithmic approach that will yield invariants more directly. The answer is a qualified yes. There is a general method, based on Cartan's method of moving frames, that is generally applicable, but finding good solutions with it still requires some ingenuity in general. Olver and co-authors have developed~\cite{Fels1999} and popularised the approach, and made it applicable to object recognition based on outline curves, see~\citet{Olver2005} for an overview, and note particularly the original application paper~\citep{Calabi1998}.  We introduce the method here and then demonstrate its use to find invariants to further groups, but for further technical details see~\citet{OlverCIT}.
